---
title: "Lab 9 - Multiple Linear Regression"
author: "Sam Csik"
output:
  html_document: default
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Use this template to follow along in Lab Week 9. Each code chunk you'll need is already created and named. 

**Lab 9 Objectives:**

- Explore multivariate data (SLO housing prices)
- Perform multiple linear regression
- Assess diagnostics
- Compare different models by AIC
- Explain model outputs
- Make a nice table of regression results
- Make predictions using a final model
- Visualize predictions with original data

###1. Load packages

- tidyverse
- stargazer

```{r packages, include = FALSE}

# a. Load packages 'tidyverse' and 'stargazer':

library(tidyverse)
library(stargazer)
```

###2. Load data (slo_homes.csv) as a df called 'homes', then filter to create a data frame called 'homes_sub' that only include homes in SLO, Arroyo Grande, Santa Maria-Orcutt, and Atascadero

```{r get_data, include = FALSE}

# a. Read in data as 'homes':

homes <- read_csv("slo_homes.csv")

# H1: as square footage increases, housing price shold increase
#H2: forclosure should sell for less than an non-foreclosed home

# b. Filter to only include cities "San Luis Obispo", "Santa Maria-Orcutt", "Atascadero", and "Arroyo Grande", and call this new subset 'homes_sub':

homes_sub <- homes %>% 
  filter(City == "San Luis Obispo" | City == "Santa Maria-Orcutt" | City == "Atascadero" | City == "Arroyo Grande")

```

###3. Go exploring (visual) + think critically about variables

*Note: It's OK to LOOK at things separately, even if you're including all in a model together!*

Example: if we want to compare distribution of housing prices by CITY (ignoring all other variables), we can do that:

```{r by_city}

# a. Calculate mean price by city
mean_by_city <- homes_sub %>% 
  group_by(City) %>% 
  summarize(
    mean = mean(Price)
  )

# b. Visualize prices by city
by_city <- ggplot(homes_sub, aes(x = Price)) +
  geom_density(aes(color = City, fill = City), alpha = 0.3) + # Note: just to show what the geom_violin shows
  theme_classic() +
  scale_x_continuous(expand = c(0,0), limits = c(0,3e6)) +
  scale_y_continuous(expand = c(0,0)) +
  labs(x = "Home Prices (USD)", y = "Density")

by_city

```

Or another question: Overall relationship between home square footage and price, separated by City? 

```{r by_sqft}

# a. Relationship between square footage and price
by_sqft <- ggplot(homes_sub, aes(x = SqFt, y = Price)) +
  geom_point(aes(color = City, pch = Status), alpha = 0.5) +
  facet_wrap(~Status)

by_sqft

# Observations here: Does relationship appear ~ linear? (YES) Anything else we can pick out re: trends, outliers, etc.? What is the general trend? Any outliers? Is there reason enough for us to omit it?

# KEEP IN MIND THAT THE MORE VARIABLES YOU INCLUDE ON A GRAPHIC, THE HARDER IT IS TO INTERPRET

```

###4. Multiple linear regression

Multiple linear regression in R follows the same syntax we've been using so far: 

    lm(y ~ x1 + x2 + x3..., data = df_name)
    
Let's try this model a couple of different ways: 

(1) Use all available variables (saturated model) 
(2) Use only SqFt as a predictor for "home size" generally (omit Bathrooms and Bedrooms), and omit PricePerSqFt (since it's derived from two other existing variables in the model)

Use summary() to view the model output and statistics.

```{r saturated}

# a. Saturated model: 

homes_lm1 <- lm(Price ~ City + Bedrooms + Bathrooms + SqFt + PricePerSqFt + Status, data = homes_sub) #finding model to find price as a functiono of....
#PricePerSqFt is perfectly derived (i.e. calculated from Price) from the output variable (Price / SqFt) so be VERY wary of this as a covariate; no random variation, sampling error, etc.

# we should be worried about using this model
# if everything about 2 houses was otherwise the same, if you increase the # of bedrooms by 1, the house price goes down by $30,792.80
# ...Forclosre has been selected as dummy variable...Status Regular (-19792.3)
# all of these things suggest that we have an overspecified model and maybe collinearity


# b. Model summary:

summary(homes_lm1)

```

The next model: Exclude price per square foot, and bedrooms/bathrooms (since these are largely explained by square footage...)

```{r subset}

# a. Updated model with variables City, SqFt, and Status:

homes_lm2 <- lm(Price ~ City + SqFt + Status, data = homes_sub)

# b. Model summary:

summary(homes_lm2)

# for every 1 sqft increase in a house, i'd expect this to go up $245.93 (makes sense)
# if we buy a house for a regular sale compared to a foreclosure sale, we'd expect to pay ~$200,000 more (makes sense)
# 0.94 p-value tells us we don't have enough evidence to suggest we short sales and foreclosure sales don't sig. differ, but regular and foreclosure sales do!
# house in Atascadero should sell for ~$202,000 less than in Arroyo Grande (reference variable)

# when we do multiple linaer regression, we use an adjusted R2 value (52.7% of housing price is explained by sqft??)

# other variables to include in model: distance from beach, school district, age of home, etc.
```

Wait...but what if I wanted everything to be with respect to a Regular sale status? Then I need to change my factor levels. We've done this before, here we'll use a different function (fct_relevel) from *forcats* package in the tidyverse. 

```{r fct_relevel}

# a. Set Status to class 'factor' (it was originally a character)

homes_sub$Status <- factor(homes_sub$Status)

# b. Check to ensure it's a factor now
#use class()

# c. Check levels:
#use levels()

# d. Reassign reference level of "Status" to "Regular":
#fct_relevel() function in forcats

homes_sub$Status <- fct_relevel(homes_sub$Status, "Regular") # can explicitly set level of columsn by writing them out in that order, otherwise they'll be listed alphabetically after the reference level

# e. Now run the regression again - same equation, but now the reference level is different (Status = Regular): 

homes_lm3 <- lm(Price ~ City + SqFt + Status, data = homes_sub)
summary(homes_lm3)

#note that now Regular doesn't show up in model summary because we've assigned this as the new reference variable
```

Interpret the coefficients and statistical outcomes above. 
Notes: 

###5. Model diagnostics

Remember, since we're concerned about *residuals* (distance that actual observations exist from model predictions), we can only evaluate some assumptions *after* running the regression. 

Then we can evaluate model diagnostics using the plot() function:

```{r diagnostics}

# a. Model diagnostics:

plot(homes_lm3)

# Residuals vs. Fitted: don't let a few data points influence your overall impressions about assumptions of heteroscedasticity (by and large, there's no trend here for our model)

# QQ Plot: data look very normally distributed

#if you were an investor, purchase house # 188 because it sold for really cheap despite all it's features
#house # 294 is celebrity status house

```

###6. Model comparison by Akaike Information Criterion

The AIC is a quantitative metric for model "optimization" that balances complexity with model fit. The best models are the ones that fit the data as well as possible, as simply as possible. Recall: lower AIC value indicates a *more optimal* balance - **BUT STATISTICS IS NO SUBSTITUTE FOR JUDGEMENT!!!**

```{r AIC}

# a. AIC values for each model

full_aic <- AIC(homes_lm1)
final_aic <- AIC(homes_lm3) # greater than full model!!??

# Answer: which would you pick? 

# still pic lm3 despite the fact that it has an AIC value greater than the full model, but our total comprehesive view of what's going on, we should definitely pick lm3 still (based on our conceptual understanding of processes)
```

###7. Regression tables with *stargazer*

```{r stargazer, results = 'asis'}

# a. Prepare a nice regression table:

stargazer(homes_lm1, homes_lm3, type = "html")

# Note: If you want to work with this in Word, save to html, open, copy and paste into Word. 

```

###8. Making predictions

Using your final selected model, predict the housing price for a range of home sizes, sale status, and city. 

The predict() function uses the following syntax:

      predict(model_name, newdata = new_data_name)
      
Defaults are to exclude the prediction SE and mean confidence interval - if you want to include, use arguments

      se.fit = TRUE
      interval = "confidence" 
      interval = "prediction"

First, you need to create a new data frame of values that contain ALL NECESSARY VARIABLES **with the same variable names AND level strings**.

```{r df_new}

# First, make a new data frame

# Note that the df_new created below has the SAME variable names and level strings as the original model data (otherwise R won't know how to use it...)
# Work through this on your own to figure out what it actually does:

df_new <- data.frame(City = rep(c("San Luis Obispo",
                                  "Santa Maria-Orcutt",
                                  "Atascadero",
                                  "Arroyo Grande"), 
                                each = 60), 
                     SqFt = rep(seq(from = 500,
                                    to = 3500, 
                                    length = 20), 
                                times = 12), 
                     Status = rep(c("Regular",
                                    "Foreclosure",
                                    "Short Sale"), 
                                  times = 12, 
                                  each = 20))

```

Make predictions for the new data using predict():

```{r predict}

# a. Make predictions using the new data frame:

price_predict <- predict(homes_lm3, newdata = df_new, se.fit = TRUE, interval = "confidence") # just gives you a vector of actual values

# b. Bind predictions to the data to make it actually useful:

predict_df <- data.frame(df_new, price_predict)

```

Then visualize it!

```{r graph, echo = FALSE}

# Create a line graph with square footage on the x-axis, predicted price on y-axis, with color dependent on City and facet_wrapped by sale status (follow along):

predict_graph <- ggplot(predict_df, aes(x = SqFt, y = fit.fit)) +
  geom_line(aes(color = City)) +
  geom_point(data = homes_sub, aes(x = SqFt, y = Price), alpha = 0.1) +
  facet_wrap(~Status)

# highest sales price = SLO, lowest for Santa Maria-Orcutt
# increase in home sale price with increase in sqft 
```

END LAB